---
title: "Reproducible Research week 2 assignment"
author: "scs"
date: "27/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Reproducible Research: Peer Assessment 1

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
```

## Loading and preprocessing the data
##point 1:
##1.1:Load the data (i.e. read.csv())
```{r}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
data <- read.csv('activity.csv')
```
## 1.2: Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
#data$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", data$interval), format='%H:%M')
```

## point2:
##2.1:What is mean total number of steps taken per day?
```{r}
x<-stepsperDay <- tapply(data$steps, data$date, sum, na.rm=TRUE)
```

##2.2: Make a histogram of the total number of steps taken each day
```{r}
qplot(x,geom = "histogram", xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

## 2.3: Calculate and report the mean and median total number of steps taken per day
```{r}
x1<- mean(x)   #mean of steps per day
x2 <- median(x)  #median of steps per day

```
* Mean: `r x1`
* Median:  `r x2`

-----
##point3:What is the average daily activity pattern?
##3.1: to make time series plot
```{r}
x3<- aggregate(x=list(meanSteps=data$steps), by=list(interval=data$interval), FUN=mean, na.rm=TRUE) #average steps per time block

ggplot(data=x3, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```


## 3.2: Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
x4<-maxsteps <- which.max(x3$meanSteps)

x5 <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", x3[x4,'interval'])  #time for max steps

```

* Max no. of  Steps at: `r x5`

----
## point4:Imputing missing values
##4.1:Calculate and report the total number of missing values in the dataset 
```{r}
missingvalues<- length(which(is.na(data$steps))) # number of missing values
```

* Number of missing values: `r missingvalues`

##4.2: Devise a strategy for filling in all of the missing values in the dataset.
#The missing values will be replaced by mean.
## 4.3:Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
data_imputed <- data
data_imputed$steps<-impute(data$steps, fun=mean) 
```

##### 4.4: Make a histogram of the total number of steps taken each day.Calculate and report the mean and median total number of steps taken per day. 
```{r}
x6<-stepsperday_imputed <- tapply(data_imputed$steps, data_imputed$date, sum)
qplot(x6,geom="histogram", xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
#calculation of mean and median of the imputed data
x7<-Meanofimputeddata <- mean(x6)
x8<-Medianofimputeddata <- median(x6)
```
* Mean (Imputed): `r x7`
* Median (Imputed):  `r x8`
Mean and median of the imputed data differ from the mean and median of

the original data (datawith out imputing). Imputed data mean and 

median are same. The impact of imputing is that it make the 

distribution of the imputed data symmetric. The mean of the original 

data was less than its median. It makes the distribution negatively skewed.

##point5: Are there differences in activity patterns between weekdays and weekends?
##5.1 :Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
data_imputed$dateType <-  ifelse(as.POSIXlt(data_imputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##5.2: Make a panel plot containing a time series plot

```{r}
averageddata_imputed <- aggregate(steps ~ interval + dateType, data=data_imputed, mean)
ggplot(averageddata_imputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```

----
